---
title: "TFM"
author: "Eloísa Toledo Iglesias"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r library }
library(tidyverse)
library(stats)
library(ggfortify)
library(cluster)
library(clusterSim)
library(factoextra)
library(fpc)
library(dbscan)
library(mclust)
library(clValid)
library(umap)
library(tsne)
library(openxlsx)
```

```{r data gene expression}

df_gen <- read_csv("data.csv") #Data

```

``` {r labels data}
df_labels <- read.csv("labels.csv") %>% as_tibble() #Labels
colnames(df_labels) <- c("samples", "class") #Change the name of the columns
df_labels$class <- as.factor(df_labels$class) #Convert to factor

samples <- df_labels$samples #Introduce the name of the samples in a variable
df_samples <- as.data.frame(samples) %>% as_tibble() #Create dataframe

table(df_labels$class) #Frequency of the cancer types
```

```{r exploratory analysis df_gen}

dim(df_gen) #Dimensions of the dataframe

count_var_types <- function(data_input) {
  types <- sapply(data_input, class)
  table_var_types <- table(types)
  return(table_var_types)
} #Sort the data by the variable type

count_var_types(df_gen) #Frequency table for each variable type
```

```{r statistics}

#Mean and median
media <- apply(df_gen[2:20532], 2, mean)
mediana <- apply(df_gen[2:20532], 2, median)
hist(media, col = rgb(0.3,0.9,0.4,0.6) , border =  rgb(0.3,0.5,0.4,0.6), 
     main = "Frequency of the average gene expression results", 
     xlab = "Mean")
hist(mediana, col = rgb(0.3,0.9,0.4,0.6) , border =  rgb(0.3,0.5,0.4,0.6), 
     main = "Frequency of the median results of the gene expression data", 
     xlab = "Median")

#Max and min
Sumary_vars <- list()
Sumary_vars$min_var  <- apply(df_gen[2:20532], 2, min)
Sumary_vars$max_var  <- apply(df_gen[2:20532], 2, max)
Sumary_vars$hist_min <- hist(Sumary_vars$min_var,breaks = 30,
                             col = rgb(1,0.6,0.4,0.6) , 
                             border =  rgb(0.3,0.5,0.4,0.6), 
                             main = "Frequency of the minimum values", 
                             xlab = "Minimum values")
Sumary_vars$hist_max <- hist(Sumary_vars$max_var,breaks = 30,
                             col = rgb(1,0.6,0.4,0.6) , 
                             border =  rgb(0.3,0.5,0.4,0.6), 
                             main = "Frequency of the maximum values", 
                             xlab = "Maximum values")

#Standard desviation
df_gen_1 <- df_gen[,-1] %>% as_tibble() #Delete sample columns
gene_variability <- apply(df_gen_1, 2, sd) #sd of each gene 
  #Data frame sd + sample column
hist(gene_variability,
     col = rgb(0.6,0.4,0.8,0.4) , 
     border =  rgb(0.3,0.5,0.4,0.6), 
     main = "Frequency of standard deviation results", 
     xlab = "Standard deviation") #histogram of sd 
table(gene_variability == 0) #Count of variables with sd = 0  
```

# SD = 0 data

```{r data sd0}
#Function to delete every variable with zero variability
sd_0 <- function(x, variability){
  gene_variability_df <- tibble(Gene = colnames(x), 
                                  Std_Deviation = variability)
  #Select genes with std_deviation higher than 0.5 (included)
  sd_no0 <- gene_variability_df %>% filter(!Std_Deviation == 0)
  df_gen2 <- x[, sd_no0$Gene] #Complete data for std_desviation != 0
  return(df_gen2)
}


df_gen2 <- sd_0(df_gen_1, gene_variability) #Complete data for std_desviation 
#!= 0, saved in environment

```

# PCA

```{r pca}

#PCA based feature extraction function

pca_filter <- function(y) {
  pca_results<- list() #List for variables names 
  pca_results$pc <- prcomp(y) #PCA for the data frame
  pca_results$varianza_explicada <- tibble(
    sdev= pca_results$pc$sdev #Create a data frame with the standard desviation
  ) %>% mutate(var_exp= sdev^2 / sum(sdev^2)*100) %>% rownames_to_column("dim") 
  #Add another variable with the explained variance
  pca_results$pca_var <- get_pca_var(pca_results$pc) #
  pca_results$pca_var$variable_importance <-
    pca_results$pca_var$contrib %>% #Add the variable contribution to the PCs 
    #to a data frame
    as.data.frame() %>%
    rownames_to_column(var = "genes") %>% #Add the variable with the gene numbers
    as_tibble() %>% #Convert to tibble
    gather("dim", "contrib", -genes) %>% #Gather contrib with dimension 
    mutate(dim = gsub("Dim.", "", dim), ) %>% #Delete "Dim." to leave only numbers
    left_join(pca_results$varianza_explicada) %>% #Join to "varianza_explicada" 
    #data frame
    mutate(var_ponderation = contrib * var_exp/100) #Calculate the variables 
    #importance, having into account the contribution of the variable to each PC  
    #and the PCs' importance. It is divided by 100 because two percentages are 
    #being multiplied

  pca_results$pca_var$gen_importance <-pca_results$pca_var$variable_importance %>%
    group_by(genes) %>% #Group by genes 
    summarise(var_index = sum(var_ponderation,na.rm = TRUE)) %>% #Sum the weighted 
    #variance
    arrange(desc(var_index)) %>% #Change the order to create an index from higher
    #explained variance to lower
    mutate(cumsum_idx= cumsum(var_index)) %>% #Calculate the cumulative sum for the
    #explained variance
  mutate(selection= ifelse(cumsum_idx<95,"yes","no")) #Create a variable that 
  #identifies as "yes" the genes of the index until the 95% of the explained 
  #variance is reached

  pca_results$pca_var$summary_selection <- pca_results$pca_var$gen_importance %>% 
    group_by(selection)%>%count() #Group the genes by selection and count them
  pca_results$pca_var$genes_selection <- pca_results$pca_var$gen_importance %>%
    filter(selection =="yes") #Filter those genes with selection == yes
  pca_results$pca_var$genes_selection_names <- pca_results$pca_var$genes_selection$genes
  #Create a variable with the names of the selected genes
  df_gen2 <- y[, pca_results$pca_var$genes_selection_names] %>%
    as_tibble() #Filter the initial data frame with the selected genes
  return(df_gen2)
}

set.seed(127)
df_gen2_pca <- pca_filter(df_gen2) #Data filtering by PCA 

```

##kmeans

```{r kmeans number clusters pca}

# Elbow method
elbow_kmeans_df_gen2_pca <- fviz_nbclust(df_gen2_pca, kmeans, method = "wss")

#Silhouette method
silhouette_kmeans_df_gen2_pca <- fviz_nbclust(df_gen2_pca, kmeans, 
                                               method='silhouette')

#Gap statistic method
set.seed(127)

gap_stat_kmeans <- clusGap(df_gen2_pca, FUN = kmeans, nstart = 10,
                    K.max = 10, B = 2)
fviz_gap_stat(gap_stat_kmeans)
```

```{r kmeans pca}
#Algorithm implementation
set.seed(127)

kmeans_3 <- kmeans(df_gen2_pca, 3, nstart = 10)
plot_kmeans_3 <- fviz_cluster(kmeans_3, data = df_gen2_pca)

kmeans_6 <- kmeans(df_gen2_pca, 6, nstart = 10)
plot_kmeans_6 <- fviz_cluster(kmeans_6, data = df_gen2_pca)

kmeans_8 <- kmeans(df_gen2_pca, 8, nstart = 10)
plot_kmeans_8 <- fviz_cluster(kmeans_8, data = df_gen2_pca)
```

```{r kmeans pca k 5}
#Algorithm implementation
set.seed(127)

kmeans_5 <- kmeans(df_gen2_pca, 5, nstart = 10)
plot_kmeans_5 <- fviz_cluster(kmeans_5, data = df_gen2_pca)

```

# K-Medoids Clustering (PAM)

```{r pam number clusters pca}

# Elbow method
elbow_pam_df_gen2_pca <- fviz_nbclust(df_gen2_pca, pam, method = "wss")
elbow_pam_df_gen2_pca

#Silhouette method
silhouette_pam_df_gen2_pca <- fviz_nbclust(df_gen2_pca, pam, 
                                               method='silhouette')
#Gap statistic method
set.seed(127)

gap_stat <- clusGap(df_gen2_pca, FUN = pam, nstart = 10,
                    K.max = 10, B = 2)
fviz_gap_stat(gap_stat)
```

```{r pam pca}
#Algorithms implementation
set.seed(127)

pam_3 <- pam(df_gen2_pca, 3, nstart = 10)
plot_pam_3 <- fviz_cluster(pam_3, data = df_gen2_pca)

pam_5 <- pam(df_gen2_pca, 5, nstart = 10)
plot_pam_5 <- fviz_cluster(pam_5, data = df_gen2_pca)

pam_6 <- pam(df_gen2_pca, 6, nstart = 10)
plot_pam_6 <- fviz_cluster(pam_6, data = df_gen2_pca)

pam_8 <- pam(df_gen2_pca, 8, nstart = 10)
plot_pam_8 <- fviz_cluster(pam_8, data = df_gen2_pca)
```

#CLARA

```{r clara number clusters pca}
# Elbow method
elbow_clara_df_gen2_pca <- fviz_nbclust(df_gen2_pca, clara, method = "wss")

#Silhouette method
silhouette_pam_df_gen2_pca <- fviz_nbclust(df_gen2_pca, pam, 
                                               method='silhouette')
#Gap statistic method
set.seed(127)

gap_stat <- clusGap(df_gen2_pca, FUN = pam, nstart = 10,
                    K.max = 10, B = 2)
fviz_gap_stat(gap_stat)
```

```{r clara pca}
#Algorithms implementation
set.seed(127)

clara_3 <- clara(df_gen2_pca, 3, samples = 10, pamLike = T)
plot_clara_3 <- fviz_cluster(clara_3, data = df_gen2_pca)

clara_5 <- clara(df_gen2_pca, 5, samples = 10, pamLike = T)
plot_clara_5 <- fviz_cluster(clara_5, data = df_gen2_pca)

clara_6 <- clara(df_gen2_pca, 6, samples = 10, pamLike = T)
plot_clara_6 <- fviz_cluster(clara_6, data = df_gen2_pca)

clara_7 <- clara(df_gen2_pca, 7, samples = 10, pamLike = T)
plot_clara_7 <- fviz_cluster(clara_7, data = df_gen2_pca)

clara_8 <- clara(df_gen2_pca, 8, samples = 10, pamLike = T)
plot_clara_8 <- fviz_cluster(clara_8, data = df_gen2_pca)
```

# DBSCAN

```{r DBSCAN}

knn_dist_df_gen2_pca <- kNNdistplot(df_gen2_pca,k=5) %>%
  abline(h=200,lty = 2) #Selecting the eps value

#Algorithm implementation

set.seed(127)

dbscan_df_gen2_pca_1 <- dbscan(df_gen2_pca, eps=174, minPts = 5, borderPoints = FALSE)
dbscan_df_gen2_pca
plot_dbscan_df_gen2_pca_1 <- fviz_cluster(dbscan_df_gen2_pca_1, df_gen2_pca, geom = "point")

dbscan_df_gen2_pca_2 <- dbscan(df_gen2_pca, eps=171, minPts = 5, borderPoints = FALSE)
dbscan_df_gen2_pca_2
plot_dbscan_df_gen2_pca_2 <- fviz_cluster(dbscan_df_gen2_pca_2, df_gen2_pca, geom = "point")

dbscan_df_gen2_pca_3 <- dbscan(df_gen2_pca, eps=165, minPts = 5, borderPoints = FALSE)
dbscan_df_gen2_pca_3
plot_dbscan_df_gen2_pca_3 <- fviz_cluster(dbscan_df_gen2_pca_3, df_gen2_pca, geom = "point")

dbscan_df_gen2_pca_5 <- dbscan(df_gen2_pca, eps=161, minPts = 5, borderPoints = FALSE)
dbscan_df_gen2_pca_5
plot_dbscan_df_gen2_pca_5 <- fviz_cluster(dbscan_df_gen2_pca_5, df_gen2_pca, geom = "point")

```

## Hierarchical

```{r hierarchical pca}
#Agglomerative
#Algorithms implementation
agglo_df_gen2_pca_complete <- agnes(df_gen2_pca, method = "complete")
agglo_df_gen2_pca_average <- agnes(df_gen2_pca, method = "average")
agglo_df_gen2_pca_single <- agnes(df_gen2_pca, method = "single")
agglo_df_gen2_pca_ward <- agnes(df_gen2_pca, method = "ward")

#Agglomerative coefficient
agglo_df_gen2_pca_complete$ac
agglo_df_gen2_pca_average$ac
agglo_df_gen2_pca_single$ac
agglo_df_gen2_pca_ward$ac

#Algorithm representation
fviz_dend(agglo_df_gen2_pca_ward, k = 5, # Cut in four groups
          cex = 0.5, # label size
          color_labels_by_k = TRUE, # color labels by groups
          rect = TRUE # Add rectangle around groups
          )

agglo_df_gen2_pca_ward_clusters <-cutree(agglo_df_gen2_pca_ward, k = 5) #Clusters

#Divisive
#Algorithms implementation
div_df_gen2__pca <- diana(df_gen2_pca)
div_df_gen2__pca$dcs #Divisive coefficient

```

## Gaussian mixture

```{r gaussian mixture pca}

gaussian_df_gen2_pca <- Mclust(df_gen2_pca) #Algorithms implementation
summary(gaussian_df_gen2_pca) #summary
 
fviz_mclust(gaussian_df_gen2_pca, "BIC", palette = "jco") #Model selection plot

fviz_mclust(gaussian_df_gen2_pca, "classification", geom = "point", 
            pointsize = 1.5, palette = "jco") #Representation

gaussian_df_gen2_pca_clusters <- gaussian_df_gen2_pca$classification #Clusters
```

## Internal evaluation

```{r Davies - Bouldin pca}

DB_results<- list()
dist_df_gen2_pca <- dist(df_gen2_pca) #Distance matrix, used for calculations
#if centrotypes = "medoids"

#kmeans

## k = 3
DB_results$kmeans_3 <- index.DB(df_gen2_pca, kmeans_3$cluster, centrotypes="centroids")
DB_results$kmeans_3$DB

## k = 5
DB_results$kmeans_5 <- index.DB(df_gen2_pca, kmeans_5$cluster, centrotypes="centroids")
DB_results$kmeans_5$DB

## k = 6
DB_results$kmeans_6 <- index.DB(df_gen2_pca, kmeans_6$cluster, centrotypes="centroids")
DB_results$kmeans_6$DB

## k = 8
DB_results$kmeans_8 <- index.DB(df_gen2_pca, kmeans_8$cluster, centrotypes="centroids")
DB_results$kmeans_8$DB

#PAM

## k = 3
DB_results$pam_3 <- index.DB(df_gen2_pca, pam_3$clustering, dist_df_gen2_pca,
                             centrotypes="medoids")
DB_results$pam_3$DB

## k = 5
DB_results$pam_5 <- index.DB(df_gen2_pca, pam_5$clustering, dist_df_gen2_pca, 
                             centrotypes="medoids")
DB_results$pam_5$DB

## k = 6
DB_results$pam_6 <- index.DB(df_gen2_pca, pam_6$clustering, dist_df_gen2_pca, 
                             centrotypes="medoids")
DB_results$pam_6$DB
## k = 8
DB_results$pam_8 <- index.DB(df_gen2_pca, pam_8$clustering, dist_df_gen2_pca, 
                             centrotypes="medoids")
DB_results$pam_8$DB

#CLARA

## k = 3
DB_results$clara_3 <- index.DB(df_gen2_pca, clara_3$clustering, dist_df_gen2_pca, 
                               centrotypes="medoids")
DB_results$clara_3$DB

## k = 5
DB_results$clara_5 <- index.DB(df_gen2_pca, clara_5$clustering,dist_df_gen2_pca, 
                               centrotypes="medoids")
DB_results$clara_5$DB

## k = 6
DB_results$clara_6 <- index.DB(df_gen2_pca, clara_6$clustering, dist_df_gen2_pca, 
                               centrotypes="medoids")
DB_results$clara_6$DB

## k = 7
DB_results$clara_7 <- index.DB(df_gen2_pca, clara_7$clustering, dist_df_gen2_pca, 
                               centrotypes="medoids")
DB_results$clara_6$DB

## k = 8
DB_results$clara_8 <- index.DB(df_gen2_pca, clara_8$clustering, dist_df_gen2_pca, 
                               centrotypes="medoids")
DB_results$clara_8$DB

```

```{r Calinski - Harabasz pca}

calinski_results <- list()

#kmeans

## k = 3
calinski_results$kmeans_3 <- calinhara(df_gen2_pca, kmeans_3$cluster)

## k = 5
calinski_results$kmeans_5 <- calinhara(df_gen2_pca, kmeans_5$cluster)

## k = 6
calinski_results$kmeans_6 <- calinhara(df_gen2_pca, kmeans_6$cluster)

## k = 8
calinski_results$kmeans_8 <- calinhara(df_gen2_pca, kmeans_8$cluster)

#PAM

## k = 3
calinski_results$pam_3 <- calinhara(df_gen2_pca, pam_3$clustering)

## k = 5
calinski_results$pam_5 <- calinhara(df_gen2_pca, pam_5$clustering)

## k = 6
calinski_results$pam_6 <- calinhara(df_gen2_pca, pam_6$clustering)

## k = 8
calinski_results$pam_8 <- calinhara(df_gen2_pca, pam_8$clustering)

#CLARA

## k = 3
calinski_results$clara_3 <- calinhara(df_gen2_pca, clara_3$clustering)

## k = 5
calinski_results$clara_5 <- calinhara(df_gen2_pca, clara_5$clustering)

## k = 6
calinski_results$clara_6 <- calinhara(df_gen2_pca, clara_6$clustering)

## k = 7
calinski_results$clara_7 <- calinhara(df_gen2_pca, clara_7$clustering)

## k = 8
calinski_results$clara_8 <- calinhara(df_gen2_pca, clara_8$clustering)

#DBScan

## eps 174
calinski_results_pca$dbscan_174 <- calinhara(df_gen2_pca, dbscan_df_gen2_pca_1$cluster)

## eps 171
calinski_results_pca$dbscan_171 <- calinhara(df_gen2_pca, dbscan_df_gen2_pca_2$cluster)

## eps 165
calinski_results_pca$dbscan_165 <- calinhara(df_gen2_pca, dbscan_df_gen2_pca_3$cluster)

## eps 161
calinski_results_pca$dbscan_161 <- calinhara(df_gen2_pca, dbscan_df_gen2_pca_5$cluster)

# Hierarchical
calinski_results$agglo_pca <- calinhara(df_gen2_pca, agglo_df_gen2_pca_ward_clusters)

# Gaussian mixture
calinski_results_pca$gaussian <- calinhara(df_gen2_pca, gaussian_df_gen2_pca$classification)
```

```{r clValid pca}

df_gen2_pca <- data.frame(df_gen2_pca) #Convert to dataframe

#Internal

clValid_results_pca <- clValid(df_gen2_pca, nClust = 3:9,
              clMethods = c("kmeans", "pam", "clara", "hierarchical", "model"),
              validation = "internal", maxitems = 801)
summary(clValid_results_pca) #Summary of the internal parameters

#Connectivity dbscan

# eps 174
connectivity(distance = dist_df_gen2_pca, dbscan_df_gen2_pca_1$cluster, 
             Data = df_gen2_pca)
# eps 171
connectivity(distance = dist_df_gen2_pca, dbscan_df_gen2_pca_2$cluster, 
             Data = df_gen2_pca)
# eps 165
connectivity(distance = dist_df_gen2_pca, dbscan_df_gen2_pca_3$cluster, 
             Data = df_gen2_pca)
# eps 161
connectivity(distance = dist_df_gen2_pca, dbscan_df_gen2_pca_5$cluster, 
             Data = df_gen2_pca)
#Dunn dbscan

# eps 174
dunn(dist_df_gen2_pca, dbscan_df_gen2_pca_1$cluster, df_gen2_pca)

# eps 171
dunn(dist_df_gen2_pca, dbscan_df_gen2_pca_2$cluster, df_gen2_pca)

# eps 165
dunn(dist_df_gen2_pca, dbscan_df_gen2_pca_3$cluster, df_gen2_pca)

# eps 161
dunn(dist_df_gen2_pca, dbscan_df_gen2_pca_5$cluster, df_gen2_pca)

#Silhouette dbscan

# eps 174
summary(silhouette(dbscan_df_gen2_pca_1$cluster, dist_df_gen2_pca))

# eps 171
summary(silhouette(dbscan_df_gen2_pca_2$cluster, dist_df_gen2_pca))

# eps 165
summary(silhouette(dbscan_df_gen2_pca_3$cluster, dist_df_gen2_pca))

# eps 161
summary(silhouette(dbscan_df_gen2_pca_5$cluster, dist_df_gen2_pca))
```

## Classification assessment

```{r classification function}

classification_function <- function(df1, df2) {
  coincidencias <- 0 #Variable initialization
  no_coincidencias <- 0 #Variable initialization
  check <- 0 #Variable initialization
  check_result <- data.frame() #Data frame initialization
  # Loop para comparar
  for(nivel in levels(df2$class)){ #For each cluster in the algorithm result
    for (level in levels(df1$class)){ #For each group in the original data
      coincidencias <- 0 #Restart variable 
      no_coincidencias <- 0 #Restart variable
      check <- 0 #Restart variable
      df1_loop <- df1[df1$class == level, ] #Select the df1 data for that level
      df2_loop <- df2[df2$class == nivel, ] #Select the df2 data for that "nivel"
      for (i in 1:nrow(df1_loop)) { #Create a loop that add 1 to coincidencias 
        #if the line i in df1 is equal to one of the lines of df2
        for (j in 1:nrow(df2_loop)){
          if (identical(df1_loop$samples[i], df2_loop$samples[j])) {
            coincidencias <- coincidencias + 1
          } 
        }  
        if (check == coincidencias){ #If in the previous loop there was no
          #addition, then add 1 to coincidencias 
          no_coincidencias <- no_coincidencias + 1
        }
        check <- coincidencias #Equal coincidencias to check to enter the next 
        #iteration of the loop
      }
    #Create data frame con los resultados
    check_result_0 <- data.frame("labels_data_cancer" = level,
                                 "labels_cluster_algorithm" = nivel,
                                 "coincidences" = coincidencias,
                                 "no_coincidences" = no_coincidencias)
    check_result <- rbind(check_result, check_result_0)
    }
  } 
  return(check_result)
}

```

``` {r kmeans classification evaluation pca}
#Classification function implementation

## k = 3
labels_kmeans_3_pca <- data.frame("samples" = samples, "class" = kmeans_3$cluster)
labels_kmeans_3_pca$class <- as.factor(labels_kmeans_3_pca$class)

write.xlsx(classification_function(df_labels, labels_kmeans_3_pca), 
           "classification/pca_kmeans_3.xlsx")

## k = 5
labels_kmeans_5_pca <- data.frame("samples" = samples, "class" = kmeans_5$cluster)
labels_kmeans_5_pca$class <- as.factor(labels_kmeans_5_pca$class)

write.xlsx(classification_function(df_labels, labels_kmeans_5_pca), 
           "classification/pca_kmeans_5.xlsx")

## k = 6
labels_kmeans_6_pca <- data.frame("samples" = samples, "class" = kmeans_6$cluster)
labels_kmeans_6_pca$class <- as.factor(labels_kmeans_6_pca$class)

write.xlsx(classification_function(df_labels, labels_kmeans_6_pca), 
           "classification/pca_kmeans_6.xlsx")

## k = 8
labels_kmeans_8_pca <- data.frame("samples" = samples, "class" = kmeans_8$cluster)
labels_kmeans_8_pca$class <- as.factor(labels_kmeans_8_pca$class)

write.xlsx(classification_function(df_labels, labels_kmeans_8_pca), 
           "classification/pca_kmeans_8.xlsx")

```

``` {r pam classification evaluation pca}
#Classification function implementation

## k = 3
labels_pam_3_pca <- data.frame("samples" = samples, "class" = pam_3$cluster)
labels_pam_3_pca$class <- as.factor(labels_pam_3_pca$class)

write.xlsx(classification_function(df_labels, labels_pam_3_pca), 
           "classification/pca_pam_3.xlsx")

## k = 5
labels_pam_5_pca <- data.frame("samples" = samples, "class" = pam_5$cluster)
labels_pam_5_pca$class <- as.factor(labels_pam_5_pca$class)

write.xlsx(classification_function(df_labels, labels_pam_5_pca), 
           "classification/pca_pam_5.xlsx")

## k = 6
labels_pam_6_pca <- data.frame("samples" = samples, "class" = pam_6$cluster)
labels_pam_6_pca$class <- as.factor(labels_pam_6_pca$class)

write.xlsx(classification_function(df_labels, labels_pam_6_pca), 
           "classification/pca_pam_6.xlsx")

## k = 8
labels_pam_8_pca <- data.frame("samples" = samples, "class" = pam_8$cluster)
labels_pam_8_pca$class <- as.factor(labels_pam_8_pca$class)

write.xlsx(classification_function(df_labels, labels_pam_8_pca), 
           "classification/pca_pam_8.xlsx")
```

``` {r clara classification evaluation pca}
#Classification function implementation

## k = 3
labels_clara_3_pca <- data.frame("samples" = samples, "class" = clara_3$cluster)
labels_clara_3_pca$class <- as.factor(labels_clara_3_pca$class)

write.xlsx(classification_function(df_labels, labels_clara_3_pca), 
           "classification/pca_clara_3.xlsx")

## k = 5
labels_clara_5_pca <- data.frame("samples" = samples, "class" = clara_5$cluster)
labels_clara_5_pca$class <- as.factor(labels_clara_5_pca$class)

write.xlsx(classification_function(df_labels, labels_clara_5_pca), 
           "classification/pca_clara_5.xlsx")

## k = 6
labels_clara_6_pca <- data.frame("samples" = samples, "class" = clara_6$cluster)
labels_clara_6_pca$class <- as.factor(labels_clara_6_pca$class)

write.xlsx(classification_function(df_labels, labels_clara_6_pca), 
           "classification/pca_clara_6.xlsx")

## k = 7
labels_clara_7_pca <- data.frame("samples" = samples, "class" = clara_7$cluster)
labels_clara_7_pca$class <- as.factor(labels_clara_7_pca$class)

write.xlsx(classification_function(df_labels, labels_clara_7_pca), 
           "classification/pca_clara_7.xlsx")

## k = 8
labels_clara_8_pca <- data.frame("samples" = samples, "class" = clara_8$cluster)
labels_clara_8_pca$class <- as.factor(labels_clara_8_pca$class)

write.xlsx(classification_function(df_labels, labels_clara_8_pca), 
           "classification/pca_clara_8.xlsx")
```

``` {r DBSCAN classification evaluation pca}
#Classification function implementation

# eps 161
labels_dbscan_161_pca <- data.frame("samples" = samples, 
                                    "class" = dbscan_df_gen2_pca_5$cluster)
labels_dbscan_161_pca$class <- as.factor(labels_dbscan_161_pca$class)

write.xlsx(classification_function(df_labels, labels_dbscan_161_pca), 
           "classification/pca_dbscan_161.xlsx")

# eps 165
labels_dbscan_165_pca <- data.frame("samples" = samples, 
                                    "class" = dbscan_df_gen2_pca_3$cluster)
labels_dbscan_165_pca$class <- as.factor(labels_dbscan_165_pca$class)

write.xlsx(classification_function(df_labels, labels_dbscan_165_pca), 
           "classification/pca_dbscan_165.xlsx")

# eps 171

labels_dbscan_171_pca <- data.frame("samples" = samples, 
                                    "class" = dbscan_df_gen2_pca_2$cluster)
labels_dbscan_171_pca$class <- as.factor(labels_dbscan_171_pca$class)

write.xlsx(classification_function(df_labels, labels_dbscan_171_pca), 
           "classification/pca_dbscan_171.xlsx")

# eps 174

labels_dbscan_174_pca <- data.frame("samples" = samples, 
                                    "class" = dbscan_df_gen2_pca_1$cluster)
labels_dbscan_174_pca$class <- as.factor(labels_dbscan_174_pca$class)

write.xlsx(classification_function(df_labels, labels_dbscan_174_pca), 
           "classification/pca_dbscan_175.xlsx")
```

``` {r hierarchical classification evaluation pca}
#Classification function implementation
labels_agglo_pca <- data.frame("samples" = samples, 
                               "class" = agglo_df_gen2_pca_ward_clusters)
labels_agglo_pca$class <- as.factor(labels_agglo_pca$class)

write.xlsx(classification_function(df_labels, labels_agglo_pca), 
           "classification/pca_agglo.xlsx")

```

``` {r gaussian classification evaluation pca}
#Classification function implementation
labels_gaussian_pca <- data.frame("samples" = samples, 
                                  "class" = gaussian_df_gen2_pca$classification)
labels_gaussian_pca$class <- as.factor(labels_gaussian_pca$class)

write.xlsx(classification_function(df_labels, labels_gaussian_pca), 
           "classification/pca_gaussian.xlsx")

```

# PCA_800

```{r pca 800}

#PCA based feature extraction function for selecting the 800 most relevant v
#variables

pca_filter <- function(y) {
  pca_results<- list() #List for variables names 
  pca_results$pc <- prcomp(y) #PCA for the data frame
  pca_results$varianza_explicada <- tibble(
    sdev= pca_results$pc$sdev #Create a data frame with the standard desviation
  ) %>% mutate(var_exp= sdev^2 / sum(sdev^2)*100) %>% rownames_to_column("dim") 
  #Add another variable with the explained variance
  pca_results$pca_var <- get_pca_var(pca_results$pc) 
  pca_results$pca_var$variable_importance <-
    pca_results$pca_var$contrib %>% #Add the variable contribution to the PCs 
    #to a data frame
    as.data.frame() %>%
    rownames_to_column(var = "genes") %>% #Add the variable with the gene numbers
    as_tibble() %>% #Convert to tibble
    gather("dim", "contrib", -genes) %>% #Gather contrib with dimension 
    mutate(dim = gsub("Dim.", "", dim), ) %>% #Delete "Dim." to leave only numbers
    left_join(pca_results$varianza_explicada) %>% #Join to "varianza_explicada" 
    #data frame
    mutate(var_ponderation = contrib * var_exp/100) #Calculate the variables 
    #importance, having into account the contribution of the variable to each PC  
    #and the PCs' importance. It is divided by 100 because two percentages are 
    #being multiplied

  pca_results$pca_var$gen_importance <-pca_results$pca_var$variable_importance %>%
    group_by(genes) %>% #Group by genes 
    summarise(var_index = sum(var_ponderation,na.rm = TRUE)) %>% #Sum the weighted 
    #variance
    arrange(desc(var_index)) #Change the order to create an index from higher
    #explained variance to lower
    
  pca_results$pca_var$genes_selection <- pca_results$pca_var$gen_importance[1:800,]
    #Filter those genes with selection == yes
  pca_results$pca_var$genes_selection_names <- pca_results$pca_var$genes_selection$genes
  #Create a variable with the names of the selected genes
  df_gen2_800 <- y[, pca_results$pca_var$genes_selection_names] %>%
    as_tibble() #Filter the initial data frame with the selected genes
  return(df_gen2_800)
}

set.seed(127)
df_gen2_pca_800 <- pca_filter(df_gen2) #Data filtering by PCA 

```

## kmeans

```{r pca_800 number clusters kmeans}
#Elbow method
elbow_kmeans_df_gen2_pca_800 <- fviz_nbclust(df_gen2_pca_800, kmeans, method = "wss")

#Silhoette method
silhouette_kmeans_df_gen2_pca_800 <- fviz_nbclust(df_gen2_pca_800, kmeans, 
                                               method='silhouette')
#Gap statistic
set.seed(127)

gap_stat_kmeans_pca_800 <- clusGap(df_gen2_pca_800, FUN = kmeans, nstart = 10,
                    K.max = 10, B = 2)
fviz_gap_stat(gap_stat_kmeans_pca_800)
```

```{r kmeans pca_800}
#Algorithm implementation
set.seed(127) 

kmeans_5_pca_800 <- kmeans(df_gen2_pca_800, 5, nstart = 10)
plot_kmeans_5_pca_800 <- fviz_cluster(kmeans_5_pca_800, data = df_gen2_pca_800)

kmeans_6_pca_800 <- kmeans(df_gen2_pca_800, 6, nstart = 10)
plot_kmeans_6_pca_800 <- fviz_cluster(kmeans_6_pca_800, data = df_gen2_pca_800)

kmeans_9_pca_800 <- kmeans(df_gen2_pca_800, 9, nstart = 10)
plot_kmeans_9_pca_800 <- fviz_cluster(kmeans_9_pca_800, data = df_gen2_pca_800)

```

## PAM

```{r pca_800 number clusters pam}
#Elbow method
elbow_pam_df_gen2_pca_800 <- fviz_nbclust(df_gen2_pca_800, pam, method = "wss")

#Silhoette method
silhouette_pam_df_gen2_pca_800 <- fviz_nbclust(df_gen2_pca_800, pam, 
                                               method='silhouette')
#Gap statistic
set.seed(127)

gap_stat_pam_pca_800 <- clusGap(df_gen2_pca_800, FUN = pam, nstart = 10,
                    K.max = 10, B = 2)
fviz_gap_stat(gap_stat_pam_pca_800)
```

```{r pam pca_800}
#Algorithm implementation
set.seed(127)

pam_5_pca_800 <- pam(df_gen2_pca_800, 5, nstart = 10)
plot_pam_5_pca_800  <- fviz_cluster(pam_5_pca_800, data = df_gen2_pca_800)

pam_6_pca_800  <- pam(df_gen2_pca_800, 6, nstart = 10)
plot_pam_6_pca_800  <- fviz_cluster(pam_6_pca_800, data = df_gen2_pca_800)

pam_9_pca_800 <- pam(df_gen2_pca_800, 9, nstart = 10)
plot_pam_9_pca_800  <- fviz_cluster(pam_9_pca_800, data = df_gen2_pca_800)
```

## CLARA

```{r pca_800 number clusters clara}
#Elbow method
elbow_clara_df_gen2_pca_800 <- fviz_nbclust(df_gen2_pca_800, clara, method = "wss")

#Silhoette method
silhouette_clara_df_gen2_pca_800 <- fviz_nbclust(df_gen2_pca_800, clara, 
                                               method='silhouette')
#Gap statistic
set.seed(127)

gap_stat_clara_pca_800 <- fviz_nbclust(df_gen2_pca_800, FUNcluster = clara, 
                               method = "gap_stat", k.max = 10, nboot = 2)

plot(gap_stat_clara_pca_800)
```

```{r clara pca_800}
#Algorithm implementation
set.seed(127)

clara_5_pca_800 <- clara(df_gen2_pca_800, 5, samples = 10, pamLike = T)
plot_clara_5_pca_800 <- fviz_cluster(clara_5_pca_800, data = df_gen2_pca_800)

clara_6_pca_800 <- clara(df_gen2_pca_800, 6, samples = 10, pamLike = T)
plot_clara_6_pca_800 <- fviz_cluster(clara_6_pca_800, data = df_gen2_pca_800)

clara_9_pca_800 <- clara(df_gen2_pca_800, 9, samples = 10, pamLike = T)
plot_clara_9_pca_800 <- fviz_cluster(clara_9_pca_800, data = df_gen2_pca_800)
```

## DBSCAN

```{r DBSCAN pca_800}

knn_dist_df_gen2_pca_800_k5 <- kNNdistplot(df_gen2_pca_800,k=5) %>%
  abline(h=85,lty = 2) #Selecting the eps value

knn_dist_df_gen2_pca_800_k6 <- kNNdistplot(df_gen2_pca_800,k=6) %>%
  abline(h=85,lty = 2) #Selecting the eps value

#Algorithm implementation

set.seed(127)

dbscan_df_gen2_pca_800_1 <- dbscan(df_gen2_pca_800, eps=85, minPts = 5, borderPoints = FALSE)
dbscan_df_gen2_pca_800_1
plot_dbscan_df_gen2_pca_800_1 <- fviz_cluster(dbscan_df_gen2_pca_800_1, df_gen2_pca_800, geom = "point")

dbscan_df_gen2_pca_800_2 <- dbscan(df_gen2_pca_800, eps=80, minPts = 5, borderPoints = FALSE)
dbscan_df_gen2_pca_800_2
plot_dbscan_df_gen2_pca_800_2 <- fviz_cluster(dbscan_df_gen2_pca_800_2, df_gen2_pca_800, geom = "point")

dbscan_df_gen2_pca_800_3 <- dbscan(df_gen2_pca_800, eps=97, minPts = 5, borderPoints = FALSE)
dbscan_df_gen2_pca_800_3
plot_dbscan_df_gen2_pca_800_3 <- fviz_cluster(dbscan_df_gen2_pca_800_3, df_gen2_pca_800, geom = "point")

```

## Hierarchical

```{r hierarchical pca_800}
#Agglomerative
#Algorithm implementation
agglo_df_gen2_pca_800_complete <- agnes(df_gen2_pca_800, method = "complete")
agglo_df_gen2_pca_800_average <- agnes(df_gen2_pca_800, method = "average")
agglo_df_gen2_pca_800_single <- agnes(df_gen2_pca_800, method = "single")
agglo_df_gen2_pca_800_ward <- agnes(df_gen2_pca_800, method = "ward")

#Agglomerative coefficients
agglo_df_gen2_pca_800_complete$ac
agglo_df_gen2_pca_800_average$ac
agglo_df_gen2_pca_800_single$ac
agglo_df_gen2_pca_800_ward$ac

fviz_dend(agglo_df_gen2_pca_800_ward, k = 6, # Cut in four groups
          cex = 0.5, # label size
          color_labels_by_k = TRUE, # color labels by groups
          rect = TRUE # Add rectangle around groups
          ) #Representation

#Clusters
agglo_df_gen2_pca_800_ward_clusters <-cutree(agglo_df_gen2_pca_800_ward, k = 6) 

#Divisive
#Algorithm implementation
div_df_gen2__pca_800 <- diana(df_gen2_pca_800)
div_df_gen2__pca_800$dc #Divisive coefficient
```

## Gaussian mixture

```{r gaussian mixture pca_800}

gaussian_df_gen2_pca_800 <- Mclust(df_gen2_pca_800) #Algorithm implementation
summary(gaussian_df_gen2_pca_800) #summary
 
fviz_mclust(gaussian_df_gen2_pca_800, "BIC", palette = "jco") #Model selection plot


fviz_mclust(gaussian_df_gen2_pca_800, "classification", geom = "point", 
            pointsize = 1.5, palette = "jco") #Representation

gaussian_df_gen2_pca_800_clusters <- gaussian_df_gen2_pca_800$classification #Clusters
```

## Internal evaluation

```{r Davies - Bouldin pca_800}

DB_results_pca_800<- list()
dist_df_gen2_pca_800 <- dist(df_gen2_pca_800) #Distance matrix, used for calculations
#if centrotypes = "medoids"

#kmeans

## k = 5
DB_results_pca_800$kmeans_5 <- index.DB(df_gen2_pca_800, kmeans_5_pca_800$cluster, 
                                     centrotypes="centroids")
DB_results_pca_800$kmeans_5$DB

## k = 6
DB_results_pca_800$kmeans_6 <- index.DB(df_gen2_pca_800, kmeans_6_pca_800$cluster, 
                                     centrotypes="centroids")
DB_results_pca_800$kmeans_6$DB

## k = 9
DB_results_pca_800$kmeans_9 <- index.DB(df_gen2_pca_800, kmeans_9_pca_800$cluster, 
                                     centrotypes="centroids")
DB_results_pca_800$kmeans_9$DB

#PAM

## k = 5
DB_results_pca_800$pam_5_pca_800 <- index.DB(df_gen2_pca_800, pam_5_pca_800$clustering, 
                                       dist_df_gen2_pca_800, centrotypes="medoids")
DB_results_pca_800$pam_5_pca_800$DB

## k = 6
DB_results_pca_800$pam_6_pca_800 <- index.DB(df_gen2_pca_800, pam_6_pca_800$clustering, 
                             dist_df_gen2_pca_800, centrotypes="medoids")
DB_results_pca_800$pam_6_pca_800$DB

## k = 9
DB_results_pca_800$pam_9_pca_800 <- index.DB(df_gen2_pca_800, pam_9_pca_800$clustering, 
                             dist_df_gen2_pca_800, centrotypes="medoids")
DB_results_pca_800$pam_9_pca_800$DB

#CLARA

## k = 5
DB_results_pca_800$clara_5_pca_800 <- index.DB(df_gen2_pca_800,
                                               clara_5_pca_800$clustering,
                                               dist_df_gen2_pca_800,
                                               centrotypes="medoids")
DB_results_pca_800$clara_5_pca_800$DB

## k = 6
DB_results_pca_800$clara_6_pca_800 <- index.DB(df_gen2_pca_800,
                                               clara_6_pca_800$clustering,
                                               dist_df_gen2_pca_800,
                                               centrotypes="medoids")
DB_results_pca_800$clara_6_pca_800$DB

## k = 9
DB_results_pca_800$clara_9_pca_800 <- index.DB(df_gen2_pca_800,
                                               clara_9_pca_800$clustering,
                                               dist_df_gen2_pca_800,
                                               centrotypes="medoids")
DB_results_pca_800$clara_9_pca_800$DB

```

```{r Calinski - Harabasz pca_800}

calinski_results_pca_800<- list()

#kmeans

## k = 5
calinski_results_pca_800$kmeans_5_pca_800 <- calinhara(df_gen2_pca_800, kmeans_5_pca_800$cluster)

## k = 6
calinski_results_pca_800$kmeans_6_pca_800 <- calinhara(df_gen2_pca_800, kmeans_6_pca_800$cluster)

## k = 9
calinski_results_pca_800$kmeans_9_pca_800 <- calinhara(df_gen2_pca_800, kmeans_9_pca_800$cluster)

#PAM

## k = 5
calinski_results_pca_800$pam_5_pca_800 <- calinhara(df_gen2_pca_800, pam_5_pca_800$clustering)

## k = 6
calinski_results_pca_800$pam_6_pca_800 <- calinhara(df_gen2_pca_800, pam_6_pca_800$clustering)

## k = 9
calinski_results_pca_800$pam_9_pca_800 <- calinhara(df_gen2_pca_800, pam_9_pca_800$clustering)

#CLARA

## k = 5
calinski_results_pca_800$clara_5_pca_800 <- calinhara(df_gen2_pca_800, clara_5_pca_800$clustering)

## k = 6
calinski_results_pca_800$clara_6_pca_800 <- calinhara(df_gen2_pca_800, clara_6_pca_800$clustering)

## k = 9
calinski_results_pca_800$clara_9_pca_800 <- calinhara(df_gen2_pca_800, clara_9_pca_800$clustering)

#DBScan

## eps 85
calinski_results_pca_800$dbscan_85_pca_800 <- calinhara(df_gen2_pca_800, dbscan_df_gen2_pca_800_1$cluster)

## eps 80
calinski_results_pca_800$dbscan_80_pca_800 <- calinhara(df_gen2_pca_800, dbscan_df_gen2_pca_800_2$cluster)

## eps 97
calinski_results_pca_800$dbscan_97_pca_800 <- calinhara(df_gen2_pca_800, dbscan_df_gen2_pca_800_3$cluster)

# Hierarchical
calinski_results_pca_800$agglo_pca_800 <- calinhara(df_gen2_pca_800, agglo_df_gen2_pca_800_ward_clusters)

# Gaussian mixture
calinski_results_pca_800$gaussian_pca_800 <- calinhara(df_gen2_pca_800, gaussian_df_gen2_pca_800$classification)
```

```{r clValid pca_800}

df_gen2_pca_800 <- data.frame(df_gen2_pca_800) #Convert to dataframe

#Internal

clValid_results_pca_800 <- clValid(df_gen2_pca_800, nClust = 5:9,
              clMethods = c("kmeans", "pam", "clara", "agnes", "model"),
              validation = "internal", maxitems = 801) #Internal evaluation
summary(clValid_results_pca_800) #Summary of the internal parameters

#Connectivity dbscan

# eps 85
connectivity(distance = dist_df_gen2_pca_800, dbscan_df_gen2_pca_800_1$cluster, 
             Data = df_gen2_pca_800)
# eps 80
connectivity(distance = dist_df_gen2_pca_800, dbscan_df_gen2_pca_800_2$cluster, 
             Data = df_gen2_pca_800)
# eps 97
connectivity(distance = dist_df_gen2_pca_800, dbscan_df_gen2_pca_800_3$cluster, 
             Data = df_gen2_pca_800)
#Dunn dbscan

# eps 85
dunn(dist_df_gen2_pca_800, dbscan_df_gen2_pca_800_1$cluster, df_gen2_pca_800)

# eps 80
dunn(dist_df_gen2_pca_800, dbscan_df_gen2_pca_800_2$cluster, df_gen2_pca_800)

# eps 97
dunn(dist_df_gen2_pca_800, dbscan_df_gen2_pca_800_3$cluster, df_gen2_pca_800)

#Silhouette dbscan

# eps 85
summary(silhouette(dbscan_df_gen2_pca_800_1$cluster, dist_df_gen2_pca_800))

# eps 80
summary(silhouette(dbscan_df_gen2_pca_800_2$cluster, dist_df_gen2_pca_800))

# eps 97
summary(silhouette(dbscan_df_gen2_pca_800_3$cluster, dist_df_gen2_pca_800))
```


## Classification assessment

``` {r kmeans classification evaluation pca_800}
#Classification function implementation
## k = 5
labels_kmeans_5_pca_800 <- data.frame("samples" = samples, "class" = kmeans_5_pca_800$cluster)
labels_kmeans_5_pca_800$class <- as.factor(labels_kmeans_5_pca_800$class)
table(kmeans_9_pca_800$cluster)

write.xlsx(classification_function(df_labels, labels_kmeans_5_pca_800), "classification/pca_800_kmeans_5.xlsx")

## k = 6
labels_kmeans_6_pca_800 <- data.frame("samples" = samples, "class" = kmeans_6_pca_800$cluster)
labels_kmeans_6_pca_800$class <- as.factor(labels_kmeans_6_pca_800$class)

write.xlsx(classification_function(df_labels, labels_kmeans_6_pca_800), "classification/pca_800_kmeans_6.xlsx")

## k = 9
labels_kmeans_9_pca_800 <- data.frame("samples" = samples, "class" = kmeans_9_pca_800$cluster)
labels_kmeans_9_pca_800$class <- as.factor(labels_kmeans_9_pca_800$class)

write.xlsx(classification_function(df_labels, labels_kmeans_9_pca_800), "classification/pca_800_kmeans_9.xlsx")

```

``` {r pam classification evaluation pca_800}
#Classification function implementation
## k = 5
labels_pam_5_pca_800 <- data.frame("samples" = samples, "class" = pam_5_pca_800$cluster)
labels_pam_5_pca_800$class <- as.factor(labels_pam_5_pca_800$class)

write.xlsx(classification_function(df_labels, labels_pam_5_pca_800), "classification/pca_800_pam_5.xlsx")

## k = 6
labels_pam_6_pca_800 <- data.frame("samples" = samples, "class" = pam_6_pca_800$cluster)
labels_pam_6_pca_800$class <- as.factor(labels_pam_6_pca_800$class)

write.xlsx(classification_function(df_labels, labels_pam_6_pca_800), "classification/pca_800_pam_6.xlsx")

## k = 9
labels_pam_9_pca_800 <- data.frame("samples" = samples, "class" = pam_9_pca_800$cluster)
labels_pam_9_pca_800$class <- as.factor(labels_pam_9_pca_800$class)

write.xlsx(classification_function(df_labels, labels_pam_9_pca_800), "classification/pca_800_pam_9.xlsx")
```

``` {r clara classification evaluation pca_800}
#Classification function implementation
## k = 5
labels_clara_5_pca_800 <- data.frame("samples" = samples, "class" = clara_5_pca_800$cluster)
labels_clara_5_pca_800$class <- as.factor(labels_clara_5_pca_800$class)

write.xlsx(classification_function(df_labels, labels_clara_5_pca_800), "classification/pca_800_clara_5.xlsx")

## k = 6
labels_clara_6_pca_800 <- data.frame("samples" = samples, "class" = clara_6_pca_800$cluster)
labels_clara_6_pca_800$class <- as.factor(labels_clara_6_pca_800$class)

write.xlsx(classification_function(df_labels, labels_clara_6_pca_800), "classification/pca_800_clara_6.xlsx")

## k = 9
labels_clara_9_pca_800 <- data.frame("samples" = samples, "class" = clara_9_pca_800$cluster)
labels_clara_9_pca_800$class <- as.factor(labels_clara_9_pca_800$class)

write.xlsx(classification_function(df_labels, labels_clara_9_pca_800), "classification/pca_800_clara_9.xlsx")
```

``` {r DBSCAN classification evaluation pca_800}
#Classification function implementation
# eps 85
labels_dbscan_85_pca_800 <- data.frame("samples" = samples, "class" = dbscan_df_gen2_pca_800_1$cluster)
labels_dbscan_85_pca_800$class <- as.factor(labels_dbscan_85_pca_800$class)

write.xlsx(classification_function(df_labels, labels_dbscan_85_pca_800), "classification/pca_800_dbscan_85.xlsx")

# eps 80
labels_dbscan_80_pca_800 <- data.frame("samples" = samples, "class" = dbscan_df_gen2_pca_800_2$cluster)
labels_dbscan_80_pca_800$class <- as.factor(labels_dbscan_80_pca_800$class)

write.xlsx(classification_function(df_labels, labels_dbscan_80_pca_800), "classification/pca_800_dbscan_80.xlsx")

# eps 97

labels_dbscan_97_pca_800 <- data.frame("samples" = samples, "class" = dbscan_df_gen2_pca_800_3$cluster)
labels_dbscan_97_pca_800$class <- as.factor(labels_dbscan_97_pca_800$class)

write.xlsx(classification_function(df_labels, labels_dbscan_97_pca_800), "classification/pca_800_dbscan_97.xlsx")
```

``` {r hierarchical classification evaluation pca_800}
#Classification function implementation
labels_agglo_pca_800 <- data.frame("samples" = samples, "class" = agglo_df_gen2_pca_800_ward_clusters)
labels_agglo_pca_800$class <- as.factor(labels_agglo_pca_800$class)

write.xlsx(classification_function(df_labels, labels_agglo_pca_800), "classification/pca_800_agglo.xlsx")

```

``` {r gaussian classification evaluation pca_800}
#Classification function implementation
labels_gaussian_pca_800 <- data.frame("samples" = samples, "class" = gaussian_df_gen2_pca_800$classification)
labels_gaussian_pca_800$class <- as.factor(labels_gaussian_pca_800$class)

write.xlsx(classification_function(df_labels, labels_gaussian_pca_800), "classification/pca_800_gaussian.xlsx")

```

# UMAP

```{r umap}
set.seed(127)

gen2_umap <- umap(df_gen2) #UMAP technique
df_gen2_umap <- as.data.frame(gen2_umap["layout"]) %>% as_tibble() #Convert to
#dataframe and tibble

plot(df_gen2_umap) #Representation
```

## kmeans

```{r umap number clusters kmeans}
#Elbow method
elbow_kmeans_df_gen2_umap <- fviz_nbclust(df_gen2_umap, kmeans, method = "wss")

#Silhoette method
silhouette_kmeans_df_gen2_umap <- fviz_nbclust(df_gen2_umap, kmeans, 
                                               method='silhouette')
#Gap statistic
set.seed(127)

gap_stat_kmeans_umap <- clusGap(df_gen2_umap, FUN = kmeans, nstart = 10,
                    K.max = 10, B = 2)
fviz_gap_stat(gap_stat_kmeans_umap)
```

```{r kmeans umap}
#Algorithm implementation
set.seed(127)

kmeans_5_umap <- kmeans(df_gen2_umap, 5, nstart = 10)
plot_kmeans_5_umap <- fviz_cluster(kmeans_5_umap, data = df_gen2_umap)

kmeans_6_umap <- kmeans(df_gen2_umap, 6, nstart = 10)
plot_kmeans_6_umap <- fviz_cluster(kmeans_6_umap, data = df_gen2_umap)

kmeans_7_umap <- kmeans(df_gen2_umap, 7, nstart = 10)
plot_kmeans_7_umap <- fviz_cluster(kmeans_7_umap, data = df_gen2_umap)

```
## PAM

```{r umap number clusters pam}
#Elbow method
elbow_pam_df_gen2_umap <- fviz_nbclust(df_gen2_umap, pam, method = "wss")

#Silhoette method
silhouette_pam_df_gen2_umap <- fviz_nbclust(df_gen2_umap, pam, 
                                               method='silhouette')
#Gap statistic
set.seed(127)

gap_stat_pam_umap <- clusGap(df_gen2_umap, FUN = pam, nstart = 10,
                    K.max = 10, B = 2)
fviz_gap_stat(gap_stat_pam_umap)
```

```{r pam umap}
#Algorithm implementation
set.seed(127)

pam_5_umap <- pam(df_gen2_umap, 5, nstart = 10)
plot_pam_5_umap  <- fviz_cluster(pam_5_umap, data = df_gen2_umap)

pam_6_umap  <- pam(df_gen2_umap, 6, nstart = 10)
plot_pam_6_umap  <- fviz_cluster(pam_6_umap, data = df_gen2_umap)

pam_7_umap <- pam(df_gen2_umap, 7, nstart = 10)
plot_pam_7_umap  <- fviz_cluster(pam_7_umap, data = df_gen2_umap)

pam_8_umap <- pam(df_gen2_umap, 8, nstart = 10)
plot_pam_8_umap  <- fviz_cluster(pam_8_umap, data = df_gen2_umap)
```

## CLARA

```{r umap number clusters clara}
#Elbow method
elbow_clara_df_gen2_umap <- fviz_nbclust(df_gen2_umap, clara, method = "wss")

#Silhoette method
silhouette_clara_df_gen2_umap <- fviz_nbclust(df_gen2_umap, clara, 
                                               method='silhouette')
#Gap statistic
set.seed(127)

gap_stat_clara_umap <- fviz_nbclust(df_gen2_umap, FUNcluster = clara, 
                               method = "gap_stat", k.max = 10, nboot = 2)

plot(gap_stat_clara_umap)
```

```{r clara umap}
#Algorithm implementation
set.seed(127)

clara_5_umap <- clara(df_gen2_umap, 5, samples = 10, pamLike = T)
plot_clara_5_umap <- fviz_cluster(clara_5_umap, data = df_gen2_umap)

clara_6_umap <- clara(df_gen2_umap, 6, samples = 10, pamLike = T)
plot_clara_6_umap <- fviz_cluster(clara_6_umap, data = df_gen2_umap)

clara_7_umap <- clara(df_gen2_umap, 7, samples = 10, pamLike = T)
plot_clara_7_umap <- fviz_cluster(clara_7_umap, data = df_gen2_umap)

clara_8_umap <- clara(df_gen2_umap, 8, samples = 10, pamLike = T)
plot_clara_8_umap <- fviz_cluster(clara_8_umap, data = df_gen2_umap)
```

## DBSCAN

```{r DBSCAN umap}

knn_dist_df_gen2_umap_k5 <- kNNdistplot(df_gen2_umap,k=5) %>%
  abline(h=0.3,lty = 2) #Selecting the eps value

#Algorithm implementation

set.seed(127)

dbscan_df_gen2_umap_1 <- dbscan(df_gen2_umap, eps=0.3, minPts = 5, borderPoints = FALSE)
dbscan_df_gen2_umap_1
plot_dbscan_df_gen2_umap_1 <- fviz_cluster(dbscan_df_gen2_umap_1, df_gen2_umap, geom = "point")

dbscan_df_gen2_umap_2 <- dbscan(df_gen2_umap, eps=0.25, minPts = 5, borderPoints = FALSE)
dbscan_df_gen2_umap_2
plot_dbscan_df_gen2_umap_2 <- fviz_cluster(dbscan_df_gen2_umap_2, df_gen2_umap, geom = "point")

dbscan_df_gen2_umap_3 <- dbscan(df_gen2_umap, eps=0.37, minPts = 5, borderPoints = FALSE)
dbscan_df_gen2_umap_3
plot_dbscan_df_gen2_umap_3 <- fviz_cluster(dbscan_df_gen2_umap_3, df_gen2_umap, geom = "point")

```

## Hierarchical

```{r hierarchical umap}
#Agglomerative
#Algorithm implementation
agglo_df_gen2_umap_complete <- agnes(df_gen2_umap, method = "complete")
agglo_df_gen2_umap_average <- agnes(df_gen2_umap, method = "average")
agglo_df_gen2_umap_single <- agnes(df_gen2_umap, method = "single")
agglo_df_gen2_umap_ward <- agnes(df_gen2_umap, method = "ward")

#Agglomerative coefficient
agglo_df_gen2_umap_complete$ac
agglo_df_gen2_umap_average$ac
agglo_df_gen2_umap_single$ac
agglo_df_gen2_umap_ward$ac

fviz_dend(agglo_df_gen2_umap_ward, k = 5, # Cut in four groups
          cex = 0.5, # label size
          color_labels_by_k = TRUE, # color labels by groups
          rect = TRUE # Add rectangle around groups
          ) #Representation

agglo_df_gen2_umap_ward_clusters <-cutree(agglo_df_gen2_umap_ward, k = 5) #Clusters

#Divisive
div_df_gen2_umap <- diana(df_gen2_umap) #Algorithm implementation
div_df_gen2_umap$dc #Divisive coefficient
```

## Gaussian mixture

```{r gaussian mixture umap}

gaussian_df_gen2_umap <- Mclust(df_gen2_umap) #Algorithm implementation
summary(gaussian_df_gen2_umap) #summary
 
fviz_mclust(gaussian_df_gen2_umap, "BIC", palette = "jco") #Model selection plot

fviz_mclust(gaussian_df_gen2_umap, "classification", geom = "point", 
            pointsize = 1.5, palette = "jco") #Representation

gaussian_df_gen2_umap_clusters <- gaussian_df_gen2_umap$classification #Clusters
```

## Internal evaluation

```{r Davies - Bouldin umap}

DB_results_umap<- list()
dist_df_gen2_umap <- dist(df_gen2_umap) #Distance matrix, used for calculations
#if centrotypes = "medoids"

#kmeans

## k = 5
DB_results_umap$kmeans_5 <- index.DB(df_gen2_umap, kmeans_5_umap$cluster, 
                                     centrotypes="centroids")
DB_results_umap$kmeans_5$DB

## k = 6
DB_results_umap$kmeans_6 <- index.DB(df_gen2_umap, kmeans_6_umap$cluster, 
                                     centrotypes="centroids")
DB_results_umap$kmeans_6$DB

## k = 8
DB_results_umap$kmeans_7 <- index.DB(df_gen2_umap, kmeans_7_umap$cluster, 
                                     centrotypes="centroids")
DB_results_umap$kmeans_7$DB

#PAM

## k = 5
DB_results_umap$pam_5_umap <- index.DB(df_gen2_umap, pam_5_umap$clustering, 
                                       dist_df_gen2_umap, centrotypes="medoids")
DB_results_umap$pam_5_umap$DB

## k = 6
DB_results_umap$pam_6_umap <- index.DB(df_gen2_umap, pam_6_umap$clustering, 
                             dist_df_gen2_umap, centrotypes="medoids")
DB_results_umap$pam_6_umap$DB

## k = 7
DB_results_umap$pam_7_umap <- index.DB(df_gen2_umap, pam_7_umap$clustering, 
                             dist_df_gen2_umap, centrotypes="medoids")
DB_results_umap$pam_7_umap$DB

## k = 8
DB_results_umap$pam_8_umap <- index.DB(df_gen2_umap, pam_8_umap$clustering, 
                             dist_df_gen2_umap, centrotypes="medoids")
DB_results_umap$pam_8_umap$DB

#CLARA

## k = 5
DB_results_umap$clara_5_umap <- index.DB(df_gen2_umap, clara_5_umap$clustering,
                               dist_df_gen2_umap, centrotypes="medoids")
DB_results_umap$clara_5_umap$DB

## k = 6
DB_results_umap$clara_6_umap <- index.DB(df_gen2_umap, clara_6_umap$clustering, 
                               dist_df_gen2_umap, centrotypes="medoids")
DB_results_umap$clara_6_umap$DB

## k = 7
DB_results_umap$clara_7_umap <- index.DB(df_gen2_umap, clara_7_umap$clustering, 
                               dist_df_gen2_umap, centrotypes="medoids")
DB_results_umap$clara_7_umap$DB

## k = 8
DB_results_umap$clara_8_umap <- index.DB(df_gen2_umap, clara_8_umap$clustering, 
                               dist_df_gen2_umap, centrotypes="medoids")
DB_results_umap$clara_8_umap$DB

```

```{r Calinski - Harabasz umap}

calinski_results_umap<- list()

#kmeans

## k = 5
calinski_results_umap$kmeans_5_umap <- calinhara(df_gen2_umap, kmeans_5_umap$cluster)

## k = 6
calinski_results_umap$kmeans_6_umap <- calinhara(df_gen2_umap, kmeans_6_umap$cluster)

## k = 7
calinski_results_umap$kmeans_7_umap <- calinhara(df_gen2_umap, kmeans_7_umap$cluster)

#PAM

## k = 5
calinski_results_umap$pam_5_umap <- calinhara(df_gen2_umap, pam_5_umap$clustering)

## k = 6
calinski_results_umap$pam_6_umap <- calinhara(df_gen2_umap, pam_6_umap$clustering)

## k = 7
calinski_results_umap$pam_7_umap <- calinhara(df_gen2_umap, pam_7_umap$clustering)

## k = 8
calinski_results_umap$pam_8_umap <- calinhara(df_gen2_umap, pam_8_umap$clustering)

#CLARA

## k = 5
calinski_results_umap$clara_5_umap <- calinhara(df_gen2_umap, clara_5_umap$clustering)

## k = 6
calinski_results_umap$clara_6_umap <- calinhara(df_gen2_umap, clara_6_umap$clustering)

## k = 7
calinski_results_umap$clara_7_umap <- calinhara(df_gen2_umap, clara_7_umap$clustering)

## k = 8
calinski_results_umap$clara_8_umap <- calinhara(df_gen2_umap, clara_8_umap$clustering)

#DBScan

## 6 clusters
calinski_results_umap$dbscan_6_umap <- calinhara(df_gen2_umap, dbscan_df_gen2_umap_1$cluster)

## 7 clusters
calinski_results_umap$dbscan_7_umap <- calinhara(df_gen2_umap, dbscan_df_gen2_umap_2$cluster)

## 9 clusters
calinski_results_umap$dbscan_9_umap <- calinhara(df_gen2_umap, dbscan_df_gen2_umap_3$cluster)

# Hierarchical
calinski_results_umap$agglo_umap <- calinhara(df_gen2_umap, agglo_df_gen2_umap_ward_clusters)

# Gaussian mixture
calinski_results_umap$gaussian_umap <- calinhara(df_gen2_umap, gaussian_df_gen2_umap$classification)
```

```{r clValid umap}

df_gen2_umap <- data.frame(df_gen2_umap) #Convert to dataframe

#Internal

clValid_results_umap <- clValid(df_gen2_umap, nClust = 5:8,
              clMethods = c("kmeans", "pam", "clara", "agnes", "model"),
              validation = "internal", maxitems = 801) #Internal evaluation
summary(clValid_results_umap) #Summary of the internal parameters

#Connectivity dbscan

# 6 clusters
connectivity(distance = dist_df_gen2_umap, dbscan_df_gen2_umap_3$cluster, 
             Data = df_gen2_umap)
# 7 clusters
connectivity(distance = dist_df_gen2_umap, dbscan_df_gen2_umap_1$cluster, 
             Data = df_gen2_umap)
# 9 clusters
connectivity(distance = dist_df_gen2_umap, dbscan_df_gen2_umap_2$cluster, 
             Data = df_gen2_umap)
#Dunn dbscan

# 6 clusters
dunn(dist_df_gen2_umap, dbscan_df_gen2_umap_3$cluster, df_gen2_umap)

# 7 clusters
dunn(dist_df_gen2_umap, dbscan_df_gen2_umap_1$cluster, df_gen2_umap)

# 9 clusters
dunn(dist_df_gen2_umap, dbscan_df_gen2_umap_2$cluster, df_gen2_umap)

#Silhouette dbscan

# 6 clusters
summary(silhouette(dbscan_df_gen2_umap_3$cluster, dist_df_gen2_umap))

# 7 clusters
summary(silhouette(dbscan_df_gen2_umap_1$cluster, dist_df_gen2_umap))

# 9 clusters
summary(silhouette(dbscan_df_gen2_umap_2$cluster, dist_df_gen2_umap))
```

```{r stability umap}
#Stability evaluations of the best algorithms
rownames(df_gen2_umap) <- c(1:801) #Add row names
clValid_stability_umap <- clValid(data.frame(df_gen2_umap), nClust = 5,
              clMethods = c("kmeans", "pam", "clara", "hierarchical"),
              validation = "stability") #Stability evaluation
summary(clValid_stability_umap) #Summary of the stability parameters
```

## Classification assessment

``` {r kmeans classification evaluation umap}
#Classification function implementation
## k = 5
labels_kmeans_5_umap <- data.frame("samples" = samples, "class" = kmeans_5_umap$cluster)
labels_kmeans_5_umap$class <- as.factor(labels_kmeans_5_umap$class)

write.xlsx(classification_function(df_labels, labels_kmeans_5_umap), "classification/umap_kmeans_5.xlsx")

## k = 6
labels_kmeans_6_umap <- data.frame("samples" = samples, "class" = kmeans_6_umap$cluster)
labels_kmeans_6_umap$class <- as.factor(labels_kmeans_6_umap$class)

write.xlsx(classification_function(df_labels, labels_kmeans_6_umap), "classification/umap_kmeans_6.xlsx")

## k = 7
labels_kmeans_7_umap <- data.frame("samples" = samples, "class" = kmeans_7_umap$cluster)
labels_kmeans_7_umap$class <- as.factor(labels_kmeans_7_umap$class)

write.xlsx(classification_function(df_labels, labels_kmeans_7_umap), "classification/umap_kmeans_7.xlsx")

```

``` {r pam classification evaluation umap}
#Classification function implementation
## k = 5
labels_pam_5_umap <- data.frame("samples" = samples, "class" = pam_5_umap$cluster)
labels_pam_5_umap$class <- as.factor(labels_pam_5_umap$class)

write.xlsx(classification_function(df_labels, labels_pam_5_umap), "classification/umap_pam_5.xlsx")

## k = 6
labels_pam_6_umap <- data.frame("samples" = samples, "class" = pam_6_umap$cluster)
labels_pam_6_umap$class <- as.factor(labels_pam_6_umap$class)

write.xlsx(classification_function(df_labels, labels_pam_6_umap), "classification/umap_pam_6.xlsx")

## k = 7
labels_pam_7_umap <- data.frame("samples" = samples, "class" = pam_7_umap$cluster)
labels_pam_7_umap$class <- as.factor(labels_pam_7_umap$class)

write.xlsx(classification_function(df_labels, labels_pam_7_umap), "classification/umap_pam_7.xlsx")

## k = 8
labels_pam_8_umap <- data.frame("samples" = samples, "class" = pam_8_umap$cluster)
labels_pam_8_umap$class <- as.factor(labels_pam_8_umap$class)

write.xlsx(classification_function(df_labels, labels_pam_8_umap), "classification/umap_pam_8.xlsx")

```

``` {r clara classification evaluation umap}
#Classification function implementation
## k = 5
labels_clara_5_umap <- data.frame("samples" = samples, "class" = clara_5_umap$cluster)
labels_clara_5_umap$class <- as.factor(labels_clara_5_umap$class)

write.xlsx(classification_function(df_labels, labels_clara_5_umap), "classification/umap_clara_5.xlsx")

## k = 6
labels_clara_6_umap <- data.frame("samples" = samples, "class" = clara_6_umap$cluster)
labels_clara_6_umap$class <- as.factor(labels_clara_6_umap$class)

write.xlsx(classification_function(df_labels, labels_clara_6_umap), "classification/umap_clara_6.xlsx")

## k = 7
labels_clara_7_umap <- data.frame("samples" = samples, "class" = clara_7_umap$cluster)
labels_clara_7_umap$class <- as.factor(labels_clara_7_umap$class)

write.xlsx(classification_function(df_labels, labels_clara_7_umap), "classification/umap_clara_7.xlsx")

## k = 8
labels_clara_8_umap <- data.frame("samples" = samples, "class" = clara_8_umap$cluster)
labels_clara_8_umap$class <- as.factor(labels_clara_8_umap$class)

write.xlsx(classification_function(df_labels, labels_clara_8_umap), "classification/umap_clara_8.xlsx")
```

``` {r DBScan classification evaluation umap}
#Classification function implementation
# 6 clusters
labels_dbscan_6_umap <- data.frame("samples" = samples, "class" = dbscan_df_gen2_umap_3$cluster)
labels_dbscan_6_umap$class <- as.factor(labels_dbscan_6_umap$class)

write.xlsx(classification_function(df_labels, labels_dbscan_6_umap), "classification/umap_dbscan_6.xlsx")

# 7 clusters
labels_dbscan_7_umap <- data.frame("samples" = samples, "class" = dbscan_df_gen2_umap_1$cluster)
labels_dbscan_7_umap$class <- as.factor(labels_dbscan_7_umap$class)

write.xlsx(classification_function(df_labels, labels_dbscan_7_umap), "classification/umap_dbscan_7.xlsx")

#9 clusters

labels_dbscan_9_umap <- data.frame("samples" = samples, "class" = dbscan_df_gen2_umap_2$cluster)
labels_dbscan_9_umap$class <- as.factor(labels_dbscan_9_umap$class)

write.xlsx(classification_function(df_labels, labels_dbscan_9_umap), "classification/umap_dbscan_9.xlsx")
```

``` {r hierarchical classification evaluation umap}
#Classification function implementation
labels_agglo_umap <- data.frame("samples" = samples, "class" = agglo_df_gen2_umap_ward_clusters)
labels_agglo_umap$class <- as.factor(labels_agglo_umap$class)

write.xlsx(classification_function(df_labels, labels_agglo_umap), "classification/umap_agglo.xlsx")

```

``` {r gaussian classification evaluation umap}
#Classification function implementation
labels_gaussian_umap <- data.frame("samples" = samples, "class" = gaussian_df_gen2_umap$classification)
labels_gaussian_umap$class <- as.factor(labels_gaussian_umap$class)

write.xlsx(classification_function(df_labels, labels_gaussian_umap), "classification/umap_gaussian.xlsx")

```